---
title: "Digital Power Europe? The EU in a Digitalizing World"
author: "Anke Obendiek, Timo Seidl"
date: "Summer 2021"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    template: syllabus_template.tex
  word_document: default
email: 'anke.obendiek@univie.ac.at'
fontfamily: helvet
fontsize: 12pt
geometry: margin=1in
header-includes: \linespread{1.05}
office: '*online*'
officehours: 'By appointment'
classroom: '*online*/Hörsaal 41 Hauptgebäude, 1.Stock, Stiege 8'
thanks: ''
classhours: Thursday, 09:45-11:15
web: ''

---
```{r setup, include=FALSE, cache=F, message=F, warning=F, results="hide"}
knitr::opts_chunk$set(fig.path='figs/')
knitr::opts_chunk$set(cache.path='_cache/')
knitr::opts_chunk$set(warning=F, message=F)
# knitr::opts_knit$get(kable.force.latex = TRUE)
```

```{r dates, include = FALSE}
library(tidyverse)
library(lubridate)

firstday <- "2021-03-11"
    
meetings <- ymd(firstday) + c(0:18) * weeks(1)

meeting_headers <- paste0("Week ", 1:19, ", ", months(meetings), " ", day(meetings))

weekday <- "Thursday"

# one can use this like this for example Students are expected to finish readings before `r paste0(weekday, "\'s")` class session

easter_start <- "2021-03-29"

easter_end <- "2021-04-11"

easter <- interval(easter_start, easter_end)

pentecost_start <- "2021-05-22"
pentecost_end <- "2021-05-25"
pentecost <- interval(pentecost_start, pentecost_end)

holiday1 <- "2021-05-13"

holiday2 <- "2021-06-03"

holiday3 <- "2021-06-03"




```


```{r include=FALSE}
library(RefManageR)
bib <- ReadBib("literature.bib")
myopts <- BibOptions(bib.style = "authoryear", style="latex", first.inits=FALSE, max.names = 20, no.print.fields = "isbn")
```

# Course Description

In a rapidly digitalizing world, Europe has yet to find its place. Some observers, including many European officials, lament that Europe has failed to keep pace with the United States and China - the two digital superpowers. They point to the fact Europe has yet to produce its own digital champion (Spotify is neither an Amazon nor an Alibaba), and raise the specter of European irrelevance in a world where political and economic influence increasingly rest upon technological progress. Meanwhile, other observers, again including many European officials, paint a rosier picture. They point to the fact that the EU has become the de-jure and de-facto regulator of large swaths of the internet, and has established itself as the global vanguard of digital human rights (the GDPR being the most famous example in both cases). But what is the role that Europe can, does, and should play in this brave new world of digital platforms, big data and artificial intelligence? In this seminar, we probe this question more systematically, assessing the position of the EU has and the limits of this power, asking ourselves why Europe lags behind in some areas and why it leads in others, and evaluating the possibility and desirability of a third model of digitalization - parallel to or in competition with the United States's market-driven laissez faire approach as well as China's state-led autocratic model. 


# Learning Outcomes

The course aims to introduce students to central debates in the emerging literature on digitalization, and focuses particularly on the role of the EU It provides students with the conceptual tools and analytic skills to embark on theoretical or empirical projects of their own. At the end of the course, students should be able to

-	identify and describe the main challenges that digitalization poses to advanced capitalist countries in general and to the European Union in particular;

- summarize and critically assess the strengths and weakness of theoretical approaches that try make sense of Europe's differential success in the areas of technological/economic and regulatory/normative leadership;

- apply the insights they gained from these approaches to their own projects.


# Requirements

Students are required to attend classes and come prepared (i.e., having finished and thought about the readings). In addition, there will be three types of assignments that together make up the final grade.

- First, for **three** sessions of their choosing, students need to write short response papers (half a page) that reflect on the readings and end with a question for the class (25%). And remember, questions end with a question mark. 

- Second, they are required to deliver a very short input presentation (around 5-10 minutes) for **one** session as well as prepare discussion points for the class (e.g., questions, empirical examples) (25%). The former is meant to quickly summarize the main points of the reading(s) and the latter is meant to kick off and organize the discussion.^[We will see how exactly we do this in the first session, after we know how many participants there are. But the idea is that no one person should talk for more than 5 minutes.] 

- Lastly, students need to write a relatively short term paper on a topic related to the course (up to 4000 words) (50%). The paper can be theoretical or empirical and is meant to hone in on one particular question that the students can pick themselves (although they should briefly discuss this with us in advance).


# Prerequisites

Students need no prior knowledge of academic debates on digitalization to successfully participate in the course. A general interest in the topic, basic English language skills, and a broad familiarity with the European Union are sufficient.

# Course Policy

Basically, don't cheat and try to learn stuff, details follow below.

## Grading Policy

You need to submit all the required assignments to pass the course. Your final grade will be the weighted average of these assignments. What is important to us when it comes to grading are two things. First, stick to the task at hand. If your response paper is meant to end with a question, end with a question. If your presentation is meant to be 5 minutes, make it no more than 6. If your term paper is meant to discuss one question in around 4000 words, don't try to answer half a dozen in 8000 words. It's almost a dad thing to say, but these skills are important not just at a university, but in any kind of job. Second, put a bit of effort into it, or at least make it look that way. Have some decent formatting. But also try to be clear and crisp, which is often harder than writing long and convoluted sentences. Try to prepare a presentation that you yourself would like to listen to: short, clear points, that highlight anything that you found confusing or unclear. You don't need to understand everything, have read a ton of additional literature, or write in a fancy way to get a very good grade. Just stick to the task and try to make sense. 

## E-mail Policy

You can always email us if you have an idea for a term paper, if you want to learn more about a certain topic and don't know where to start, or if you have a question that you really don't want to ask in class. Please don't email us with questions that you could easily find the answer to in the syllabus or in previous emails. We might take it badly. 

This is a tough time for many, so if you feel like you struggle to cope or just need to get an outside perspective, please get in touch with the Psychological Counseling Center: https://www.studierendenberatung.at/en 

## Attendance Policy

You are required to attend each session, and we encourage you to prepare for and actively participate in them. However, if you really can't make it, just reach out to us, these things happen once or twice a term. 



\newpage

\begingroup\huge
**Course Outline**
\endgroup

\vspace{0.25cm}



##  `r meeting_headers[1]` : Introduction


*No readings for this session*
```{r, echo = FALSE, results="asis"} 
if(meetings[1]%within%easter | meetings[1]%within%pentecost | meetings[1] == holiday1 | meetings[1] == holiday2 | meetings[1] == holiday3){
  cat(noquote("*We do not have class this week due to public holidays*"))
}
``` 

# \textcolor{Blue}{\textsc{Part I: Foundations of Digitalization}}
\vspace{0.5cm}

##  `r meeting_headers[2]` : The Rise of the Platform Society

```{r, echo = FALSE, results="asis"} 
if(meetings[2]%within%easter | meetings[2]%within%pentecost | meetings[2] == holiday1 | meetings[2] == holiday2 | meetings[2] == holiday3){
  cat(noquote("*We do not have class this week due to public holidays*"))
}
``` 

van Dijck, J. (2020) ‘Governing digital societies: Private platforms, public values’, Computer Law & Security Review, 36, pp. 1--4.

Srnicek, N. (2017) Platform Capitalism. Cambridge: Polity, pp. 36--92.


##  `r meeting_headers[3]` : Artifical Intelligence & Big Data

MIT Technology Review and Oracle (2016) The Rise of Data Capital. Available at: \url{http://files.technologyreview.com/whitepapers/MIT_Oracle+Report-The_Rise_of_Data_Capital.pdf}.

MIT Work of the Future Task Force (2020) The Work of the Future: Building Better Jobs in an Age of Intelligent Machines, pp. 1--6, pp. 30--46. Available at: \url{https://workofthefuture.mit.edu/research-post/the-work-of-the-future-building-better-jobs-in-an-age-of-intelligent-machines/}.


```{r, echo = FALSE, results="asis"} 
if(meetings[3]%within%easter | meetings[3]%within%pentecost | meetings[3] == holiday1 | meetings[3] == holiday2 | meetings[3] == holiday3){
  cat(noquote("*We do not have class this week due to public holidays*"))
}
``` 


##  `r meeting_headers[4]` : Spring holiday



```{r, echo = FALSE, results="asis"} 
if(meetings[4]%within%easter | meetings[4]%within%pentecost | meetings[4] == holiday1 | meetings[4] == holiday2 | meetings[4] == holiday3){
  cat(noquote("*We do not have class this week due to public holidays*"))
}
``` 


##  `r meeting_headers[5]` : Spring holiday

```{r, echo = FALSE, results="asis"} 
if(meetings[5]%within%easter | meetings[5]%within%pentecost | meetings[5] == holiday1 | meetings[5] == holiday2 | meetings[5] == holiday3){
  cat(noquote("*We do not have class this week due to public holidays*"))
}
``` 

# \textcolor{Blue}{\textsc{Part II: The European Union in the World}}
\vspace{0.5cm}


##  `r meeting_headers[6]` :   Normative Power Europe

Manners, I. (2002) ‘Normative Power Europe: A Contradiction in Terms?’ JCMS: Journal of Common Market Studies, 40(2), pp. 235–258. 

Diez, T. (2005). Constructing the Self and Changing Others: Reconsidering `Normative Power Europe’. Millennium, 33(3), pp. 613–636. 


```{r, echo = FALSE, results="asis"} 
if(meetings[6]%within%easter | meetings[6]%within%pentecost | meetings[6] == holiday1 | meetings[6] == holiday2 | meetings[6] == holiday3){
  cat(noquote("*We do not have class this week due to public holidays*"))
}
``` 


##  `r meeting_headers[7]` : Market Power Europe

Damro, C. (2012) ‘Market Power Europe’, Journal of European Public Policy, 19(5), pp. 682–699.

Seoane, M. V. (2020) Normative market Europe? The contested governance of cyber-surveillance technologies. In: Emerging Security Technologies and EU Governance, Routledge, pp. 88–101. 

```{r, echo = FALSE, results="asis"} 
if(meetings[7]%within%easter | meetings[7]%within%pentecost | meetings[7] == holiday1 | meetings[7] == holiday2 | meetings[7] == holiday3){
  cat(noquote("*We do not have class this week due to public holidays*"))
}
``` 


##  `r meeting_headers[8]` : The Brussels Effect

Bradford, A. (2020) The Brussels effect: How the European Union rules the world. New York, NY: Oxford University Press, Chapter 5. Digital Economy, pp. xiii-xix, pp. 131-170. 


```{r, echo = FALSE, results="asis"} 
if(meetings[8]%within%easter | meetings[8]%within%pentecost | meetings[8] == holiday1 | meetings[8] == holiday2 | meetings[8] == holiday3){
  cat(noquote("*We do not have class this week due to public holidays*"))
}
``` 

# \textcolor{Blue}{\textsc{Part III: Models of Digitalization}}
\vspace{0.5cm}

##  `r meeting_headers[9]` : The United States: All hail the Market?


```{r, echo = FALSE, results="asis"} 
if(meetings[9]%within%easter | meetings[9]%within%pentecost | meetings[9] == holiday1 | meetings[9] == holiday2 | meetings[9] == holiday3){
  cat(noquote("*We do not have class this week due to public holidays*"))
}
``` 
Farrell, H., & Newman, A. L. (2021). The Janus Face of the Liberal International Information Order: When Global Institutions Are Self-Undermining. International Organization, pp. 1–26.

Newman, A. L., & Bach, D. (2004). Self-Regulatory Trajectories in the Shadow of Public Power. Resolving Digital Dilemmas in Europe and the United States. Governance, 17(3), pp. 387–413.

##  `r meeting_headers[10]` : 

```{r, echo = FALSE, results="asis"} 
if(meetings[10]%within%easter | meetings[10]%within%pentecost | meetings[10] == holiday1 | meetings[10] == holiday2 | meetings[10] == holiday3){
  cat(noquote("*We do not have class this week due to public holidays*"))
}
``` 



##  `r meeting_headers[11]` : China: State Surveillance Capitalism? 

```{r, echo = FALSE, results="asis"} 
if(meetings[11]%within%easter | meetings[11]%within%pentecost | meetings[11] == holiday1 | meetings[11] == holiday2 | meetings[11] == holiday3){
  cat(noquote("*We do not have class this week due to public holidays*"))
}
``` 
Jia, L. and Winseck, D. (2018) ‘The political economy of Chinese internet companies: Financialization, concentration, and capitalization’, International Communication Gazette, 80(1), pp. 30-–59. 

Shen, H. (2018) ‘Building a Digital Silk Road? Situating the Internet in China’s Belt and Road Initiative’, International Journal of Communication, 12, pp. 2683–-2701.


##  `r meeting_headers[12]` : The European Union: A Third Way? 

```{r, echo = FALSE, results="asis"} 
if(meetings[12]%within%easter | meetings[12]%within%pentecost | meetings[12] == holiday1 | meetings[12] == holiday2 | meetings[12] == holiday3){
  cat(noquote("*We do not have class this week due to public holidays*"))
}
``` 

Floridi, L. (2020). The Fight for Digital Sovereignty: What It Is, and Why It Matters, Especially for the EU. Philosophy & Technology, 33(3), pp. 369–378. 

Radu, R., & Chenou, J.-M. (2015). Data control and digital regulatory space (s): Towards a new European approach. Internet Policy Review, Journal on Internet Regulation, 4(2).


# \textcolor{Blue}{\textsc{Part IV: The EU's Digital Agenda}}

\vspace{0.5cm}

##  `r meeting_headers[13]` : 

```{r, echo = FALSE, results="asis"} 
if(meetings[13]%within%easter | meetings[13]%within%pentecost | meetings[13] == holiday1 | meetings[13] == holiday2 | meetings[13] == holiday3){
  cat(noquote("*We do not have class this week due to public holidays*"))
}
``` 

##  `r meeting_headers[14]` : Data Protection 

```{r, echo = FALSE, results="asis"} 
if(meetings[13]%within%easter | meetings[13]%within%pentecost | meetings[13] == holiday1 | meetings[13] == holiday2 | meetings[13] == holiday3){
  cat(noquote("*We do not have class this week due to public holidays*"))
}
``` 

Rossi, A. (2018) ‘How the Snowden Revelations Saved the EU General Data Protection Regulation’, The International Spectator, 53(4), pp. 95-111.

Farrell, H., & Newman, A. L. (2018). Linkage Politics and Complex Governance in Transatlantic Surveillance. World Politics, 70(4), 515–554.

##  `r meeting_headers[15]` : Taxation

Danescu, E. (2020). Taxing intangible assets: Issues and challenges for a digital Europe. Internet Histories, 4(2), 196–216. 

Lips, W. (2020). The EU Commission’s digital tax proposals and its cross-platform impact in the EU and the OECD. Journal of European Integration, 42(7), 975–990.

```{r, echo = FALSE, results="asis"} 
if(meetings[15]%within%easter | meetings[15]%within%pentecost | meetings[15] == holiday1 | meetings[15] == holiday2 | meetings[15] == holiday3){
  cat(noquote("*We do not have class this week due to public holidays*"))
}
``` 




##  `r meeting_headers[16]` : Policy Session: Digital Services Act/Digital Markets Act


tba

```{r, echo = FALSE, results="asis"} 
if(meetings[16]%within%easter | meetings[16]%within%pentecost | meetings[16] == holiday1 | meetings[16] == holiday2 | meetings[16] == holiday3){
  cat(noquote("*We do not have class this week due to public holidays*"))
}
``` 




