---
output:
  xaringan::moon_reader:
    includes:
      after_body: insert-logo.Rhtml
    lib_dir: libs
    css: [xaringan-themer.css]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
    seal: false
---



```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, crayon.enabled = TRUE, htmltools.preserve.raw = FALSE)

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)


library(xaringanthemer)
library(crayon)
library(tidyverse)
library(knitr)
library(ggthemes)
library(kableExtra)
library(xts)
library(tidyquant)
library(dygraphs)
library(widgetframe)
library(rDNA)
library(igraph)
library(networkD3)
library(plotly)
library(ggtext)
library(glue)
library(RColorBrewer)
library(countrycode)
library(cowplot)
library(grid)


```


```{r xaringan-themer}


xaringanthemer::style_mono_light(
  base_color = "#2A363B",
  link_color = "#FF847C",
  text_slide_number_color = "#FECEA8",
  header_font_google = google_font("Lato"),
  text_font_google   = google_font("Roboto"),
  code_font_google   = google_font("IBM Plex Mono"),
  text_font_size = "1.25rem",
  colors = c(
    red = "#f34213",
    purple = "#3e2f5b",
    orange = "#ff8811",
    green = "#99B898",
    white = "#FFFFFF"
  )
)



# colors can be used like this:This ** .red[simple] ** .white.bg - purple[demo]
# _.orange[shows]_ the colors .green[in action].
```


class: inverse, center, middle


# Digital Power Europe

## Spring/Summer Term 2021

<br>

### Anke Obendiek & Timo Seidl

### Centre for European Integration Research

---



name: contents-slide

## Overview

.pull-left[


[Session 1: Introduction](#part1)

#### Part I: Foundations of Digitalization

[Session 2: The Rise of the Platform Society](#part2)

[Session 3: Artifical Intelligence & Big Data](#part3)


#### Part II: The European Union in the World


[Session 4: Normative Power Europe](#part4)


[Session 5: Market Power Europe](#part5)

[Session 6: The Brussels Effect](#part6)

]


.pull-right[


#### Part III: Models of Digitalization

[Session 7: The US: All hail the Market?](#part7)

[Session 8: China: State Surveillance Capitalism?](#part8)

[Session 9: The European Union: A Third Way?](#part9)


#### Part IV: The EU's Digital Agenda

[Session 10: Data Protection](#part10)

[Session 11: Taxation](#part11)

[Session 12 Policy Session: DSA and DMA](#part12)
   
]

---


name: part1
class: inverse, center, middle

# Session 1: Introduction

.footnote[

[Back to the contents slide](#contents-slide)

]

---

## Views on the European Union

```{r fig.width=16, fig.height=7}

cucumbers <- "img/cucumbers.jpg"
roaming <- "img/roaming.jpg"
papertiger <- "img/papertiger.jpg"
superpower <- "img/superpower.jpg"


eu_df <- data.frame()


p <- eu_df%>%
  ggplot()+
  # geom_vline(aes(xintercept = 0, color = "red"))+
  # # geom_hline(yintercept = 0)+
  theme_void()+
  theme(plot.background = element_rect(fill = "#E9EAEB"))


ggdraw(p)+
  draw_image(cucumbers, scale = 0.15, hjust = 0.425)+
  draw_image(roaming, scale = 0.15, hjust = -0.42)+
  draw_image(papertiger, scale = 0.15, vjust = 0.42)+
  draw_image(superpower, scale = 0.15, vjust = -0.42)+
  annotation_custom(grobTree(textGrob("Bureaucratic nuisance",  hjust=1.35, vjust = 7,
                                      gp=gpar(col="#2A363B", fontsize=25, fontface="italic"))))+
  annotation_custom(grobTree(textGrob("Benign but in decline",  hjust=-.55, vjust = 7,
                                      gp=gpar(col="#2A363B", fontsize=25, fontface="italic"))))+
  annotation_custom(grobTree(textGrob("Totalitarian superstate",  hjust=1.4, vjust = -6,
                                      gp=gpar(col="#2A363B", fontsize=25, fontface="italic"))))+
  annotation_custom(grobTree(textGrob("'Brussels Effect'",  hjust=-.75, vjust = -6,
                                      gp=gpar(col="#2A363B", fontsize=25, fontface="italic"))))
```

---
