---
name: Timo
surname: Seidl
position: "Post-Doc"
address: "Centre for European Integration Research, University of Vienna"
#phone: 
www: timoseidl.com
email: "timo.seidl@univie.ac.at"
twitter: Timo_Seidl
#github: tseidl
#linkedin: 
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  vitae::moderncv:
    theme: banking
headcolor: 19448a
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```


# Research Interests


 * **Disciplinary**: Comparative Political Economy, Politics of Public Policy, Economic Sociology 
 * **Theoretical**: Ideational Approaches, Coalitional Politics, Business Power
 * **Substantive**: Digitalization, European Integration, Capitalism and Democracy
 * **Methodological**: Text as Data, Time-Series-Cross-Section, Discourse Network Analysis

<p>&nbsp;</p>

# Academic Positions & Fellowships

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Post-Doc", "2021-", "University of Vienna, Centre for European Integration Research", "",
  "Visiting Researcher", "2020", "Alexander von Humboldt Institute for Internet and Society", "",
  "Visiting Researcher", "2020", "Humboldt University Berlin", ""
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

# Education

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Doctor of Political and Social Sciences", "2021", "European University Insitute", "",
  "Master of Research", "2018", "European University Insitute", "",
  "Semester abroad", "2016", "University of Toronto", "",
  "Master of Arts, Sociology", "2017", "Goethe University of Frankfurt", "",
  "Semester abroad", "2014", "University of Oslo", "",
  "Bachelor of Arts, Social Sciences", "2015", "University of Augsburg", ""
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```


# Grants & Scholarships

```{r}
tribble(
  ~Type, ~Year, ~Desc, ~x,
  "for studies at Humboldt University","2020", "Erasmus Scholarship", "",
  "for studies at the European University Institute","2017-", "DAAD PhD Grant", "",
  "for studies at the University of Toronto",  "2016", "DAAD Grant" ,"",
  "for fieldwork in Namibia", "2015", "Young Scholar Research Grant", "",
  "for studies at the University of Oslo","2014", "Eramus Scholarship", "",
  "Merit-based Scholarship", "2013-2017", "German Academic Scholarship Foundation", ""
  ) %>% 
  detailed_entries(Type, Year, Desc, x)
```

# Teaching (Syllabi & Materials can be found [here](https://www.timoseidl.com/teaching/))

<p>&nbsp;</p>
```{r}
tibble::tribble(
  ~Type, ~Year, ~Desc, ~x,
  "B.A. Course University of Vienna","Winter 2021", "Key Controversies in European Integration", "",
  "B.A. Course University of Vienna","Winter 2021", "The Twin Transitions: The EU's Climate and Digital Policies", "",
  "M.A. Course University of Vienna","Summer 2021", "Politics of Platform Capitalism", "Evaluation result: 1.1 (1-5 scale)",
  "M.A. Course University of Vienna","Summer 2021", "Digital Power Europe", "Evaluation result: 1.5 (1-5 scale)",
  "B.A. Course University of Vienna","Summer 2021", "Policymaking in the European Union", "Evaluation result: 1.2 (1-5 scale)",
  "B.A. Course Heinrich Heine University Düsseldorf","Summer 2020", "Introduction to Comparative Political Economy", "",
  ) %>% 
  detailed_entries(Type, x, Desc, Year,  .protect = F)
```


# Professional Activities & Development

## Organizer

<p>&nbsp;</p>

```{r}
tribble(
  ~Year, ~Type, ~Desc,
  "2018-2020", "Co-Chair Political Economy Working Group, European University Institute", "",
  "2019", "Co-Organizer of the workshop What is Social Democracy and what is it good for?, European University Institute", "",
  "2018", "Co-Organizer of the panel Max Weber and the study of 21st century capitalism, general conference German Sociological Association", ""
  ) %>% 
  brief_entries(Type, Year, Desc)
```

<p>&nbsp;</p>

## Reviewer

<p>&nbsp;</p>

Comparative Political Studies (1x), Journal of Common Market Studies (1x), Industrial Relations (1x), Regulation & Governance (1x), Swiss Journal of Sociology (1x), Zeitschrift für Politikwissenschaft (Journal of Political Science) (1x)

<p>&nbsp;</p>

## Memberships

<p>&nbsp;</p>

CES, DGS, ECPR, IPSA, SASE, Young Scholar Max Weber Network

<p>&nbsp;</p>

## Methods Training

<p>&nbsp;</p>

Web Scraping and Text Mining (with Paulo Serodio), Causal Inference (with Elias Dinas), Interviewing (with Jennifer Welsh), Qualitative Comparative Analysis (with Claudius Wagemann), Network Analysis (with Paulo Serodio), Text Analysis (with Inaki Sagarzazu)

<p>&nbsp;</p>

## Additional Experiences & Training (Selection)

<p>&nbsp;</p>

```{r}
tibble::tribble(
  ~Type, ~Year, ~Desc, ~x,
    "Introduction to Teaching in Higher Education, European University Institute","2018-2019", "Teaching Training", "",
  "Oslo Summer School, with Christel Koop, one week","2018", "The Political Economy of Public Policy", "",
  "Summer school, with Daniel Haun and Peter Kappeler, four weeks",  "2015-2017", "Evolutionary Anthropology" ,"",
  "Summer school, one week", "2015", "Comparing the Cambridge School and Conceptual History", "",
  "International aid organization","2013", "Internship Humedica", ""
  ) %>% 
  detailed_entries(Type, Year, Desc, x)
```

# Skills

## Languages

<p>&nbsp;</p>

German (native), English (TOEFL: 118 of 120 points), Italian (beginner), Latin & Ancient Greek (basic reading proficiency)

<p>&nbsp;</p>

## Programming & Formatting Languages

<p>&nbsp;</p>

R, RMarkdown, Latex, Python, Stata

<p>&nbsp;</p>

## Software

<p>&nbsp;</p>

Discourse Network Analyser, MAXQDA, Gephi, Citavi



# Publications


## Thesis

<p>&nbsp;</p>

Seidl, T. (2021) Ideas, Politics, and Technological Change: Essays on the Comparative Political Economy of Digital Capitalism. Dissertation European University Institute. | [**PDF**](https://cadmus.eui.eu/handle/1814/69757).


## Peer-Reviewed Journal Articles

<p>&nbsp;</p>

* Marenco, Matteo and Seidl, Timo (2021) ‘The discursive construction of digitalization: a comparative analysis of national discourses on the digital future of work’, European Political Science Review, pp. 1–19. 
[**Article**](https://www.cambridge.org/core/journals/european-political-science-review/article/discursive-construction-of-digitalization-a-comparative-analysis-of-national-discourses-on-the-digital-future-of-work/A4B6C6E17C4EE9AD002F7E2553C24696) |
[**Working Paper**](https://doi.org/10.31235/osf.io/6ex7t) 

* Seidl, Timo (2020) ‘The Politics of Platform Capitalism. A Case Study on the Regulation of Uber in New York’. Regulation & Governance. [**Article**](https://onlinelibrary.wiley.com/doi/abs/10.1111/rego.12353) | [**Working Paper**](https://osf.io/preprints/socarxiv/up9qz/) | [**Replication Materials**](www.timoseidl.com/publication/the-politics-of-platform-capitalism/Replication_Material_Seidl_Politics_of_Platform_Capitalism.zip)

* Laurer, Moritz & Seidl, Timo (2020) ‘Regulating the European Data-Driven Economy. A Case Study on the GDPR’. Policy & Internet [**Article**](https://onlinelibrary.wiley.com/doi/abs/10.1002/poi3.246) | [**Working Paper**](https://osf.io/preprints/socarxiv/a6m8r/) | [**Replication Materials**](www.timoseidl.com/publication/regulating-the-european-data-driven-economy/Laurer_Seidl_GDPR_Replication.zip).

* Roos, Ulrich & Seidl, Timo (2015) ‘Im »Südwesten« nichts Neues? Eine Analyse der deutschen Namibiapolitik als Beitrag zur Rekonstruktion der außenpolitischen Identität des deutschen Nationalstaates’, Zeitschrift für Friedens- und Konfliktforschung, 4(2), pp. 182–225. [**Article**](https://www.timoseidl.com/publication/im-suedwesten-nichts-neues/Im%20Suedwesten%20nichts%20neues.pdf)

<p>&nbsp;</p>

## Preprints & Working Papers

<p>&nbsp;</p>

* Seidl, Timo (2021) ‘Commodification and Disruption: Theorizing Digital Capitalism’, EIF Working Paper. [Link](https://eif.univie.ac.at/downloads/workingpapers/wp2021-02.pdf)

* Nachtwey, Oliver & Seidl, Timo (2020) ‘The Solutionist Ethic and the Spirit of Digital Capitalism’. Soc ArXiv Working Paper. [**Link**](https://doi.org/10.31235/osf.io/sgjzq)

* Nachtwey, Oliver & Seidl, Timo (2017) 'Die Ethik der Solution und der Geist des digitalen Kapitalismus'. Institute for Social Research Working Paper (#11). [**Link**](http://www.ifs.uni-frankfurt.de/wp-content/uploads/IfS-WP-11.pdf)

<p>&nbsp;</p>

## Work in Progress

<p>&nbsp;</p>

* Obendiek, Anke & Seidl, Timo 'The (False) Promise of Solutionism. Epistemic Authority and the Delegation of Core State Powers in Digital Governance', Early-stage working paper.

* Schmitz, Luuk & Seidl, Timo. 'Protecting, Transforming, and Projecting the Single Market
Strategic Autonomy in the EU’s Trade and Digital Policies', Early-stage working paper.

* Seidl, Timo: Investing in the Digital Future.	The Comparative Political Economy of Digital Investment Policies. Unpublished Manuscript.


<p>&nbsp;</p>

## Book Chapters

<p>&nbsp;</p>

* Nachtwey, Oliver & Seidl, Timo (2020) ‘Ideologie und der Geist des digitalen Kapitalismus’, in Nienhüser, W. and Matiaske, W. (eds.) Ökonomie und Ideologie. (Jahrbuch Ökonomie und Gesellschaft). Weimar: Metropolis, pp. 235–265. 

* Laurer, Moritz & Seidl, Timo (2017) ‘Eine Stabilitäts- und Wettbewerbsunion der Regierungen: Zur Rekonstruktion grundlegender Handlungsregeln deutscher Eurokrisenpolitik’, in Roos, U. (ed.) Deutsche Außenpolitik: Arenen, Diskurse und grundlegende Handlungsregeln. Wiesbaden: Springer VS, pp. 13–47.

* Seidl, Timo (2017) ‘Kompatibilität statt Reduktion: Zur Idee einer Interdisziplinären Anthropologie als soziologischem Selbst- und Fremdver(un)sicherungsraum’, in Wagner, G. (ed.) Die Provokation der Reduktion: Beiträge zur Wissenschaftstheorie der Soziologie. Wiesbaden: Harrassowitz, pp. 201–232.

<p>&nbsp;</p>

## Book Reviews & Other Publications

<p>&nbsp;</p>

* Seidl, Timo: ‘Book Review: Appadurai, Arjun. and Alexander, Neta (2020) Failure: Cambridge. Polity Press’, Economic Sociology: European Electronic Newsletter. [**Link**](https://econsoc.mpifg.de/40764/06_BookReviewsEconsoc-NL_21-2_March2020.pdf)

* Laurer, Moritz & Seidl, Timo (2014) ‘Das doppelte Ziel der deutschen Eurokrisenpolitik’, Welttrends, 22(99), pp. 120–126.

<p>&nbsp;</p>

## Newspaper Articles & Blog Posts

<p>&nbsp;</p>

*  Seidl, Timo 'The Techlash in Times of Corona', HIIG Blog. [**Link**](https://www.hiig.de/en/the-techlash-in-times-of-corona/)

* Seidl, Timo (2018) Die Linke und der Geist des digitalen Kapitalismus, Oxi Magazin. 10/18, p. 5.

* Roos, Ulrich & Seidl, Timo (2015) Time to Call Things as They Are, Namibian, July 14, p. 7.

<p>&nbsp;</p>


# Presentations (Selection)

<p>&nbsp;</p>

## Academic Conferences

<p>&nbsp;</p>


* **SASE** (2021): Protecting, Transforming, and Projecting the Single Market, Investing in the Digital Future

* **CES** (2021): Protecting, Transforming, and Projecting the Single Market

* **ECPR** (2020): Investing in the Digital Future, Talkin’ bout Digitalization

* **GIG-ARTS** (2019): Regulating the European Data-Driven Economy

* **DGS** (2018): Max Weber im Silicon Valley

* **ECPR** (2018): Varieties of Platform Capitalism 

* **ÖGS** (2017): Homo Moralis vs. Homo Oeconomicus 

* **SGS** (2017): Neuer oder neuester Geist des Kapitalismus?

<p>&nbsp;</p>

## Non-Academic Conferences

<p>&nbsp;</p>

* **LABORA19**: The Politics of Platform Capitalism (2019)

# References

<p>&nbsp;</p>

**Dorothee Bohle**    
Professor of Political Sciences   
European University Institute    
Dept. of Political and Social Sciences   
dorothee.bohle@eui.eu   

<p>&nbsp;</p>

**Oliver Nachtwey**   
Professor of Social Structure Analysis   
University of Basel   
Dept. of Social Science   
oliver.nachtwey@unibas.ch

<p>&nbsp;</p>

**Anton Hemerijck**   
Professor of Political Science and Sociology   
European University Institute   
Dept. of Political and Social Sciences   
anton.hemerijck@eui.eu





 







